### 2015-05-19 Manuscript/Missing_mammals-draft.tex
Updated discussion

%Authors guidlines: http://royalsocietypublishing.org/instructions-authors
% 2500 words max (includes the title page, abstract, references, acknowledgements and figure/table legends)
% current version is above 4500. I think a big cut down can be done on the references.
% We allow a maximum of 4 displays, only 2 of which can be figures.

@@

\section{Discussion}
Our results show that even though relations among living mammals are well known in a molecular framework \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ} only little cladistic data is available.
This is primarily due to the fact that that morphological studies are mainly focusing on fossil taxa rather than living taxa \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
However, data availability varies in its structure as well as across taxonomic levels.
Higher taxonomic levels are always better sampled than lower ones (table \ref{Table_morpho_taxa_proportion} and within these taxonomic levels, the structure of available data is mostly random apart for two orders (table \ref{Table_data_structure}.

%\subsection{Data availability}
The amount of living taxa with no available cladistic data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available cladistic data (and two of the 28 orders are monospecific).
This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
Beyond this threshold, there is a noteworthy displacement of the wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002} as well as a important decrease in the conservation of clades \citep{GuillermeCooper}.
At the species level within most orders of mammals, it is therefore likely to see topological artefacts for the placement of fossil taxa.
However, the data availability seems to be less an issue at higher taxonomic level (i.e. the Genus and the Family level).
This point is important to consider from a practical point of view because of the slight discrepancy between neontological and palaeontological species.
While neontological species are described on accurately measured morphology, genetic distance, spacial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}; palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species), most palaeontological studies are using the Genus level for their smallest taxonomic OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
In this frame of palaeontological data usage, the data availability at the Genus level in living mammals should be our priority concern.

%\subsection{Available data structure}
Regardless the low level of available data, it is encouraging that for most orders (except Carnivora and Chiroptera), it is randomly distributed across the phylogeny.
Therefore, it is likely that the addition of fossil taxa within these orders will not be biased by oversampling but will probably branch near the taxa that has the most similar available data (as expected).
When only few data is available, the ideal scenario will be that the data is over-dispersed (i.e.
that there is data in at least every sub-clade) in order to maximize the possibilities of the fossil to branch in the right clade.
In the second scenario, when the data is randomly distributed we expected no special bias in the placement of the fossil \citep{GuillermeCooper}.
However, in the third scenario, when the data is clustered we expect two major biases to occur: (1) first, the fossil will not be able to branch within a clade containing no data (e.g.
Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverage}-B, and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g.
a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverage}-B.

%For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (figure ~\ref{Figure_example_coverage}-B.
On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis} will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverage}-B.

%Caveats
In this study, we treated all morphological matrices to be equal in a similar way that molecular matrices are: for example if a matrix A contains 100 characters for taxa X and Y, and a matrix B contains 50 characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix way leading to a matrix containing 150 characters for taxon X, 100 for taxon Y and 50 for taxon Z.
However, it is clear that cladistic data cannot be treated that way \citep{Brazeau2011}.
In fact, for the same taxon, some characters might overlap (e.g.
matrix A has a character coding for the shape of a particular morphological feature and matrix B as a first character coding for the presence of this morphological feature and a second one coding for its shape; in this case, these three characters are compound characters \citep{Brazeau2011}.
However, in reasonably sized matrices ($>$ 100 characters \citep{GuillermeCooper,harrisonamong-character2014} it is more likely that a number of characters are consistently conserved among the different matrices \citep[e.g.][]{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}.
A conservative approach to avoid compound characters could be to select only the latest matrix for each taxon.

%Let's code some data
Following the recommendations in \citep{GuillermeCooper}, one should code morphological characters for a maximum of living species in order to improve the topology of Total Evidence trees containing both living and fossil taxa.
Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}.

%Biology letters various stuff
\section{Ethics statement}
\section{Data accessibility statement}
All data is available and reproducible on GitHub.
\section{Authors’ contributions statement}
Conceived and designed the experiments: TG NC. Performed the experiments: TG. Analyzed the data: TG. Wrote the paper: TG NC.
\section{Acknowledgements}
Nick Matzke, April Wright, David Bapst and Graeme Lloyd.
\section{Funding statement}
This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).

>>>>

%Authors guidlines: http://royalsocietypublishing.org/instructions-authors
% 2500 words max (includes the title page, abstract, references, acknowledgements and figure/table legends)
% current version is around 3700. I think a big cut down can be done on the references.
% We allow a maximum of 4 displays, only 2 of which can be figures.

@@

\section{Discussion}
Our results show that even though relations among living mammals are well known in a molecular framework \citep[e.g.][]{FritzTree,meredithimpacts2011,May-Collado-PeerJ} only little cladistic data is available.
This is primarily due to the fact that that cladistic studies are mainly focusing on fossil taxa rather than living taxa \citep[e.g.][]{O'Leary08022013,ni2013oldest}.
However, data availability varies in its structure as well as across taxonomic levels.
Higher taxonomic levels are always better sampled than lower ones (table \ref{Table_morpho_taxa_proportion}) and within these taxonomic levels, the structure of available data is mostly random apart for two orders (table \ref{Table_data_structure})b.

%\subsection{Data availability}
The amount of living taxa with no available cladistic data was surprisingly high at the species level: 24/28 orders have less than 75\% taxa with available cladistic data (and two of the 28 orders are monospecific!).
This threshold of 75\% available data is the maximum amount of possible missing data (25\%) before having a significant effect on the topology of Total Evidence trees \citep{GuillermeCooper}.
Beyond this threshold, there is a noteworthy displacement of the wildcard taxa (\textit{sensu} \citep{kearneyfragmentary2002}) as well as a important decrease in the conservation of clades \citep{GuillermeCooper}.
At the species level within most orders of mammals, it is therefore likely to see topological artefacts for the placement of fossil taxa.
However, the data availability seems to be less an issue at higher taxonomic level (i.e. the Genus and the Family level).
This point is important to consider from a practical point of view because of the slight discrepancy between the neontological and the palaeontological concept of species.
While neontological species are described on accurately measured morphology, genetic distance, spacial distribution and even behaviour \citep[e.g.][]{kellymolecular2014}, palaeontological data can be based only on morphological, spatial and temporal data \citep[e.g.][]{ni2013oldest}.
Because of this difference between neontological species (\textit{sensu} reproductive isolates) and paleontological species (\textit{sensu} morpho-species %or "morphological clusters"?
), most palaeontological studies are using the Genus level for their smallest OTU \citep[e.g.][]{ni2013oldest,O'Leary08022013}.
In this frame of palaeontological data usage, the data availability at the Genus level in living mammals should be our priority concern.

%\subsection{Available data structure}
Regardless the low level of available data, it is encouraging that for most orders (except Carnivora and Chiroptera), data is randomly distributed across the phylogeny.
%Therefore, it is likely that the addition of fossil taxa within these orders will not be biased by oversampling but will probably branch near the taxa that has the most similar available data (as expected).
There are three different scenarii, regarding the distribution of the available data, for predicting how fossils can branch in relation to living taxa.
When only few data is available, the ideal scenario will be that it is over-dispersed (i.e. that there is data in at least every sub-clade) in order to maximize the possibilities of the fossil to branch in the right clade.
In the second scenario, when the data is randomly distributed we expected no special bias in the placement of the fossil \citep{GuillermeCooper}.
However, in the third scenario, when the data is clustered we expect two major biases to occur: (1) first, the fossil will not be able to branch within a clade containing no data (e.g. Herpestidae in Carnivora; figure \ref{Figure_missing_data_problem} and \ref{Figure_example_coverage}-B), and (2) second, the fossil will have a higher probability, at random, to branch within the clade containing most of the available data (e.g. a Carnivora fossil will have more chance to branch, randomly, to the Canidae clade than any other clade in Carnivora; figure \ref{Figure_example_coverage}-B).
This means that fossils with uncertain phylogentic affinities (\textit{Incertae sedis}) will have a higher probability of branching by chance within the most sampled clade.

%For example in Carnivora, their is no available data for Herpestidae, therefore any fossil Herpestidae will not be able to be correctly placed in the phylogeny (figure ~\ref{Figure_example_coverage}-B.
%On the other hand, any Carnivora fossil with uncertain phylogenetic affinities (\textit{Incertae sedis} will have a higher probability of branching by chance within the Canidae because they are oversampled within the Carnivora (figure ~\ref{Figure_example_coverage}-B.

%Caveats
In this study, we treated all cladistic matrices to be equal in a similar way that molecular matrices are: for example if a matrix A contains 100 characters for taxa X and Y, and a matrix B contains 50 characters for taxa X and Z, we assumed that both matrices can be combined in a supermatrix containing 150 characters for taxon X, 100 for taxon Y and 50 for taxon Z.
However, it is clear that cladistic data cannot be treated that way \citep{Brazeau2011}.
In fact, for the same taxon, some characters might overlap (e.g. matrix A has a character coding for the shape of a particular morphological feature and matrix B as a first character coding for the presence of this morphological feature and a second one coding for its shape; in this case, these three characters are compound characters \citep{Brazeau2011}).
However, in reasonably sized matrices ($>$ 100 characters \citep{GuillermeCooper,harrisonamong-character2014} it is more likely that a number of characters are consistently conserved among the different matrices (e.g. within primates literature, the character \textit{p7} - the size of the $4^{th}$ lower premolar paraconid - is consistent for more than 15 years \citep{ross1998phylogenetic,seiffert2003fossil,marivaux2005anthropoid,seiffert2005basal,bloch2007new,kay2008anatomy,silcox2008biogeographic,seiffert2009convergent,tabuce2009anthropoid,boyer2010astragalar,seiffert2010fossil,marivaux2013djebelemur,ni2013oldest}).
A conservative approach to avoid compound characters could be to select only the most recent matrix for each group.

%Let's code some data
Following the recommendations in \citep{GuillermeCooper}, one should code cladistic characters for a maximum of living species in order to improve the topology of Total Evidence trees containing both living and fossil taxa.
Since the data for living mammals is usually easily available in world wide vast natural history collections, we propose that an increased effort in coding morphological characters from living species should be done via engaging in collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}.

%Biology letters various stuff
\section{Ethics statement}
\section{Data accessibility statement}
All data is available and reproducible on GitHub.
\section{Authors’ contributions statement}
Conceived and designed the experiments: TG NC. Performed the experiments: TG. Analyzed the data: TG. Wrote the paper: TG NC.
\section{Acknowledgements}
Nick Matzke, April Wright, David Bapst and Graeme Lloyd. %Or whatever order.
\section{Funding statement}
This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
